### 2017-11-11 article/pos.tex
first draft almost done

\begin{table*}[h!]
\begin{center}
\caption{"POS tags incident and chosen.
            The official dictionary often relates tokens
            to more than one POS tag.
            For the text highlighting Plugin, for example,
            a token has to have an established tag to have
            a defined color.
            On the Chosen column, the tokens were regarded only once
            by choosing the first classification in the dictionary in ['PRE', 'VERB', 'PREPOSITION', 'PARTICLE', 'ADJECTIVE', 'NOUN', 'NUMBER'].}
\begin{tabular}{| l | c | c |}\hline
POS & All  & Chosen \\\hline
NOUN & 58  & 49 \\\hline
ADJECTIVE & 40  & 35 \\\hline
VERB & 18  & 13 \\\hline
PARTICLE & 12  & 12 \\\hline
PRE & 5  & 5 \\\hline
PREPOSITION & 5  & 5 \\\hline
NUMBER & 4  & 1 \\\hline
total & 142  & 120 \\\hline
\end{tabular}\end{center}
\end{table*}