### 2017-11-24 article/vowels.tex
andando com as estatisticas

\begin{table*}[h!]
\begin{center}
\caption{"Frequency of vowels in Toki Pona.}
\begin{tabular}{| l | c |}\hline
Vowel & freq \\\hline
a & 33.19 \\\hline
e & 17.45 \\\hline
i & 23.40 \\\hline
o & 15.32 \\\hline
u & 10.64 \\\hline
\end{tabular}\end{center}
\end{table*}