### 2017-12-15 article/pos.tex

\begin{table*}[h!]
\begin{center}
\caption{POS tags incident and chosen.
            The official dictionary often relates tokens
            to more than one POS tag.
            For the current version of the text highlighting plugin described in
            Section~\ref{shigh}, for example,
            a token has to have an established tag to have
            a defined color.
            On the 'Chosen' column, the tokens were regarded only once
            by choosing the class in the dictionary respecting the
            precedence order: PRE, VERB, PREPOSITION, PARTICLE, ADJECTIVE, NOUN, NUMBER.}\label{foobar}
\begin{tabular}{| l | c | c |}\hline
POS & All  & Chosen \\\hline
NOUN & 58  & 49 \\\hline
ADJECTIVE & 40  & 34 \\\hline
VERB & 15  & 13 \\\hline
PARTICLE & 12  & 12 \\\hline
PRE & 6  & 6 \\\hline
PREPOSITION & 5  & 5 \\\hline
NUMBER & 4  & 1 \\\hline
total & 140  & 120 \\\hline
\end{tabular}\end{center}
\end{table*}