### 2017-12-15 article/syls.tex

\begin{table*}[h!]
\begin{center}
\caption{Frequency of syllables in Toki Pona
            considering all 235 syllables of the 124 tokens,
            only the first or last syllables or only the middle
            syllable.
            In parenthesis are the count and percentage of the
            corresponding syllable. For more information and a
            complete list of syllables, see Section~\ref{sec:stat}.
            }\label{tab:freqsy}
\begin{tabular}{  l | c   c   c   c  }
rank & all  & first  & last  & middle \\\hline
1 & li (13, 5.53\%)  & a (8, 6.45\%)  & li (10, 8.06\%)  & la (2, 15.38\%) \\
2 & la (10, 4.26\%)  & o (5, 4.03\%)  & lo (6, 4.84\%)  & je (1, 7.69\%) \\
3 & ka (9, 3.83\%)  & pi (5, 4.03\%)  & na (6, 4.84\%)  & ka (1, 7.69\%) \\
4 & na (9, 3.83\%)  & ka (4, 3.23\%)  & la (5, 4.03\%)  & ke (1, 7.69\%) \\
5 & pa (9, 3.83\%)  & la (4, 3.23\%)  & ma (5, 4.03\%)  & li (1, 7.69\%) \\
6 & a (8, 3.40\%)  & pa (4, 3.23\%)  & pa (5, 4.03\%)  & lu (1, 7.69\%) \\
7 & ma (8, 3.40\%)  & se (4, 3.23\%)  & ka (4, 3.23\%)  & ma (1, 7.69\%) \\
8 & si (8, 3.40\%)  & si (4, 3.23\%)  & sa (4, 3.23\%)  & me (1, 7.69\%) \\
9 & lo (7, 2.98\%)  & su (4, 3.23\%)  & si (4, 3.23\%)  & pe (1, 7.69\%) \\
10 & pi (6, 2.55\%)  & i (3, 2.42\%)  & te (4, 3.23\%)  & ta (1, 7.69\%) \\
\end{tabular}\end{center}
\end{table*}