### 2018-07-02 pack/prv/opt/tokipona/aux/article/acm/pos.tex
terminada passagem de rotinas das tabelas do artigo para classes

\begin{table*}[h!]
\begin{center}
\caption{"POS tags incident and chosen as preferential e.g. in text synthesis.
                    The official dictionary often relates tokens
                    to more than one POS tag.
                    For the text highlighting Plugin, for example,
                    a token has to have an established tag to have
                    a defined color.
                    On the Chosen column, the tokens were regarded only once
                    by choosing the first occurrence of ['PRE', 'VERB', 'PREPOSITION', 'PARTICLE', 'ADJECTIVE', 'NOUN', 'NUMBER'] in the official dictionary.}\label{foobar}
\begin{tabular}{  l | c   c  }
POS & All  & Chosen \\\hline
NOUN & 58  & 49 \\
ADJECTIVE & 40  & 34 \\
VERB & 15  & 13 \\
PARTICLE & 12  & 12 \\
PRE & 6  & 6 \\
PREPOSITION & 5  & 5 \\
NUMBER & 4  & 1 \\\hline
total & 140  & 120 \\
\end{tabular}\end{center}
\end{table*}