### 2018-07-05 pack/prv/opt/tokipona/aux/article/acm/sample-acmsmall.tex
passado artigo p template correto p tallip (acmsmall)

\documentclass[format=acmsmall, review=false, screen=true]{acmart}
\usepackage[utf8]{inputenc}

\usepackage{booktabs} % For formal tables
\newcommand{\tttt}[1]{\texttt{#1}}
\usepackage{subcaption}
\usepackage{tipa}

\usepackage[ruled]{algorithm2e} % For algorithms
\renewcommand{\algorithmcfname}{ALGORITHM}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}

% Metadata Information
\acmJournal{TALLIP}
% \acmVolume{X}
% \acmNumber{X}
% \acmArticle{XX}
% \acmYear{2018}
% \acmMonth{X}
% \acmArticleSeq{XX}
\acmVolume{0}
\acmNumber{0}
\acmArticle{00}
\acmYear{2018}
\acmMonth{7}
\acmArticleSeq{00}

% Copyright
%\setcopyright{acmcopyright}
\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\acmDOI{0000001.0000001}

% Paper history
\received{July 2018}
% \received{March 2009}
\received[accepted]{XXX XXXX}

% Document starts
\begin{document}
% Title portion. Note the short title for running heads
\title[A Multifrequency MAC for Wireless Sensor]{Basic concepts and tools for the Toki Pona minimal and constructed language: description of the language and main issues; analysis of the vocabulary; text synthesis and syntax highlighting; Wordnet synsets}

\author{Renato Fabbri}
\authornote{This is the corresponding author}
\orcid{0000-0002-9699-629X}
\affiliation{%
  \institution{Institute of Mathematical and Computer Sciences, University of São Paulo}
  \streetaddress{Avenida Trabalhador São-carlense, 400 - Centro}
  \city{São Carlos}
  \state{SP}
  \postcode{668}
  \country{BR}}
\email{renato.fabbri@gmail.com}

\begin{abstract}
A minimal constructed language (conlang)
is useful for experiments and comfortable for making tools.
The Toki Pona (TP) conlang is minimal both in the vocabulary
(with only 14 letters and 124 lemmas)
and in the (about) 10 syntax rules.
The language is useful for being a used and somewhat established
minimal conlang with at least hundreds of fluent speakers.
This article exposes current concepts and resources
for TP,
  and makes available Python (and Vim) scripted routines for
the analysis of the language,
synthesis of texts, syntax highlighting schemes,
  and the achievement of a preliminary TP Wordnet.
Focus is on the analysis of the basic vocabulary,
as corpus analyses were found.
The synthesis is based on sentence templates,
relates to context by keeping track of used words,
and renders larger texts by using a fixed number of phonemes (e.g. for poems)
and number of sentences, words and letters (e.g. for paragraphs).
Syntax highlighting 
reflects morphosyntactic classes given in the official dictionary
and different solutions are described and implemented
in the well-established Vim text editor.
The tentative TP Wordnet is made available in three patterns
  of relations between synsets and word lemmas.
In summary, this text holds potentially novel conceptualizations about,
and tools and results in analyzing, synthesizing and syntax highlighting the
TP language.
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below.
%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010405.10010469</concept_id>
  <concept_desc>Applied computing~Arts and humanities</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010405.10010497</concept_id>
  <concept_desc>Applied computing~Document management and text processing</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10003456.10010927.10003619</concept_id>
  <concept_desc>Social and professional topics~Cultural characteristics</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc{Applied computing~Arts and humanities}
\ccsdesc[500]{Applied computing~Document management and text processing}
\ccsdesc[300]{Social and professional topics~Cultural characteristics}

%
% End generated code
%

\keywords{Constructed languages, natural language processing, syntax highlighting, wordnet, Toki Pona}

\maketitle

% The default list of authors is too long for headers.
\renewcommand{\shortauthors}{Fabbri, R.}

\input{samplebody-journals}

\end{document}